---
title: "Workplace_Wellness_Data_Cleaning_Code"
author: "Mary Hoyt Akiyama Kearns"
date: "1/5/2022"
output: html_document
---

## Running Descriptive Statistics

```{r install packages -- I had to add the CRAN URL to address an error preventing R Markdown from running}
install.packages("tidyverse", repos = "http://cran.us.r-project.org")
install.packages("skimr", repos = "http://cran.us.r-project.org")
install.packages("janitor", repos = "http://cran.us.r-project.org")
```

```{r load packages}
library(tidyverse)
library(skimr)
library(janitor)
library(lubridate)
library(ggplot2) 
```

```{r Import **converted** data into R}
work_wellness <- read.csv("WHA_2017.csv")
```

```{r Run descriptives}
head(work_wellness)
colnames(work_wellness)
str(work_wellness)
n_distinct(work_wellness$Id)
nrow(work_wellness)
```

There are 2843 unique IDs and 2843 rows of data, so one observation per ID. There are 301 variables, and quite a few variables coded as 96, 97, 98, 99 (missing and miscellaneous)

## Create a Subset of the Data Using Key Variables

```{r Create NEW dataset using ONLY variables of interest, and do a quick check of the columns}
key_work_wellness <- work_wellness %>% 
    select("Id", "Industry", "Size", "Region", "CDC_Region", "OC1", "OC3", "HPR8_1", "HPR8_1A", "HPR8_1B", "HPR8_1C", 
           "CP1", "CP2", "CP3", "WD1_1", "WD1_2", "WD2", "WD3", "WD4", "WD5", "WD7", "WL2", "WL5", "WL6", 
           "WL7", "WL8", "WL9", "WL11", "WL12", "WL14", "WL15","B1_1", "B1_2", "B1_3", "B1_4", "B1_5", 
           "B1_6", "B1_7", "B1_8", "B1_9", "B1_10", "B1_11", "B1_12", "KP3_1", "KP3_2", "KP3_3", "KP4", 
           "KP5A", "KP5C", "KP5E", "KP5F", "KP5G", "KP5H", "HPR8_2A_S") 
view(key_work_wellness)
colnames(key_work_wellness)
n_distinct(key_work_wellness$Id)
nrow(key_work_wellness)
```

There are now 2843 unique IDs and 2843 rows of data, for 54 variables.

## Rename columns to make them more user-friendly
```{r Rename the columns and check the names to make sure they stuck}
key_work_wellness <- key_work_wellness %>% rename(Num_Employees = "Size",
                            Your_Dept = "OC1",
                            Org_Type = "OC3", 
                            Stress_Manage = "HPR8_1", 
                            Stress_Manage_Type = "HPR8_1A", 
                            Stress_Manage_Who_Runs = "HPR8_1B", 
                            Stress_Manage_Particip = "HPR8_1C",
                            Health_Skill_Dev= "CP1",
                            Supportive_Environment= "CP2",
                            Integrate_Into_Org_Strucutre= "CP3",
                            Percent_Under_30 = "WD1_1", 
                            Percent_60_Or_Over = "WD1_2", 
                            Percent_Female = "WD2", 
                            Percent_Hourly = "WD3", 
                            Percent_Night_Shift = "WD4", 
                            Percent_Remote = "WD5", 
                            Percent_Turnover = "WD7", 
                            Bullying_Prevent = "WL2", 
                            Flexible_Sched = "WL5", 
                            Telework = "WL6", 
                            Family_Leave = "WL7", 
                            Parental_Leave_Paid = "WL8", 
                            Parental_Leave_Unpaid = "WL9", 
                            Childcare_Reimburse = "WL11", 
                            Childcare_OnOrOffsite ="WL12", 
                            Edlercare_Reimburse = "WL14", 
                            Eldercare_Onsite = "WL15",
                            Barrier_Interest = "B1_1", 
                            Barrier_Trust = "B1_2", 
                            Barrier_Cost = "B1_3", 
                            Barrier_Trained_Staff = "B1_4", 
                            Barrier_Time = "B1_5", 
                            Barrier_Privacy_Concerns = "B1_6", 
                            Barrier_Outcomes = "B1_7", 
                            Barrier_Qualified_Vendors = "B1_8",
                            Barrier_Sr_Managment_Support = "B1_9", 
                            Barrier_Mid_Managment_Support = "B1_10", 
                            Barrier_Legal_Concerns = "B1_11", 
                            Barrier_Physical_Space = "B1_12", 
                            Incentive_Attendance = "KP3_1", 
                            Incentive_Achieve_Health_Std = "KP3_2", 
                            Incentive_Attendance_Health_Std = "KP3_3", 
                            Incentive_Effectiveness = "KP4", 
                            Incentive_Premium_Discounts = "KP5A", 
                            Incentive_Copay_Diff = "KP5C", 
                            Incentive_Cash_Rewards = "KP5E", 
                            Incentive_Flex_Spending = "KP5F", 
                            Incentive_Paid_Leave = "KP5G", 
                            Incentive_Gifts_Prizes = "KP5H",
                            Yoga_Meditation_Room = "HPR8_2A_S")

colnames(key_work_wellness)
head(work_wellness)
```

## Convert variables from numeric to character, and some from character to numeric
In checking the new dataframe, I noticed that some columns were in the wrong format, so I ran the following code to convert them. This is my first attempt at coding this in R, so I'm guessing there is a much more parsimonious sytax, but this got the job done. Any feedback you have on streamlining this is greatly appreciated!

```{r}
# Convert "Industry"and "CDC_Region" from Integer to Character 
is.numeric(key_work_wellness$Industry)
key_work_wellness$Industry <- as.character(key_work_wellness$Industry)
is.character(key_work_wellness$Industry)

is.numeric(key_work_wellness$CDC_Region)
key_work_wellness$CDC_Region <- as.character(key_work_wellness$CDC_Region)
is.character(key_work_wellness$CDC_Region)

#Convert data from Character to Integer
is.character(key_work_wellness$Percent_Under_30)
suppressWarnings((key_work_wellness$Percent_Under_30 <- as.numeric(key_work_wellness$Percent_Under_30)))
is.numeric(key_work_wellness$Percent_Under_30)

is.character(key_work_wellness$Percent_60_Or_Over)
suppressWarnings(key_work_wellness$Percent_60_Or_Over <- as.numeric(key_work_wellness$Percent_60_Or_Over))
is.numeric(key_work_wellness$Percent_60_Or_Over)

is.character(key_work_wellness$Percent_Female)
suppressWarnings((key_work_wellness$Percent_Female <- as.numeric(key_work_wellness$Percent_Female)))
is.numeric(key_work_wellness$Percent_Female)

is.character(key_work_wellness$Percent_Hourly)
suppressWarnings((key_work_wellness$Percent_Hourly <- as.numeric(key_work_wellness$Percent_Hourly)))
is.numeric(key_work_wellness$Percent_Hourly)

is.character(key_work_wellness$Percent_Night_Shift)
suppressWarnings(key_work_wellness$Percent_Night_Shift <- as.numeric(key_work_wellness$Percent_Night_Shift))
is.numeric(key_work_wellness$Percent_Night_Shift)

is.character(key_work_wellness$Percent_Remote)
suppressWarnings(key_work_wellness$Percent_Remote <- as.numeric(key_work_wellness$Percent_Remote))
is.numeric(key_work_wellness$Percent_Remote)

is.character(key_work_wellness$Percent_Turnover)
suppressWarnings(key_work_wellness$Percent_Turnover <- as.numeric(key_work_wellness$Percent_Turnover))
is.numeric(key_work_wellness$Percent_Turnover)

is.character(key_work_wellness$Barrier_Interest)
suppressWarnings(key_work_wellness$Barrier_Interest <- as.numeric(key_work_wellness$Barrier_Interest))
is.numeric(key_work_wellness$Barrier_Interest)

is.character(key_work_wellness$Barrier_Trust)
suppressWarnings(key_work_wellness$Barrier_Trust <- as.numeric(key_work_wellness$Barrier_Trust))
is.numeric(key_work_wellness$Barrier_Trust)

is.character(key_work_wellness$Barrier_Cost)
suppressWarnings(key_work_wellness$Barrier_Cost <- as.numeric(key_work_wellness$Barrier_Cost))
is.numeric(key_work_wellness$Barrier_Cost)

is.character(key_work_wellness$Barrier_Trained_Staff)
suppressWarnings(key_work_wellness$Barrier_Trained_Staff <- as.numeric(key_work_wellness$Barrier_Trained_Staff))
is.numeric(key_work_wellness$Barrier_Trained_Staff)

is.character(key_work_wellness$Barrier_Time)
suppressWarnings(key_work_wellness$Barrier_Time <- as.numeric(key_work_wellness$Barrier_Time))
is.numeric(key_work_wellness$Barrier_Time)

is.character(key_work_wellness$Barrier_Privacy_Concerns)
suppressWarnings(key_work_wellness$Barrier_Privacy_Concerns <- as.numeric(key_work_wellness$Barrier_Privacy_Concerns))
is.numeric(key_work_wellness$Barrier_Privacy_Concerns)

is.character(key_work_wellness$Barrier_Outcomes)
suppressWarnings(key_work_wellness$Barrier_Outcomes <- as.numeric(key_work_wellness$Barrier_Outcomes))
is.numeric(key_work_wellness$Barrier_Outcomes)

is.character(key_work_wellness$Barrier_Qualified_Vendors)
suppressWarnings(key_work_wellness$Barrier_Qualified_Vendors <- as.numeric(key_work_wellness$Barrier_Qualified_Vendors))
is.numeric(key_work_wellness$Barrier_Qualified_Vendors)

is.character(key_work_wellness$Barrier_Sr_Managment_Support)
suppressWarnings(key_work_wellness$Barrier_Sr_Managment_Support <- as.numeric(key_work_wellness$Barrier_Sr_Managment_Support))
is.numeric(key_work_wellness$Barrier_Sr_Managment_Support)

is.character(key_work_wellness$Barrier_Mid_Managment_Support)
suppressWarnings(key_work_wellness$Barrier_Mid_Managment_Support <- as.numeric(key_work_wellness$Barrier_Mid_Managment_Support))
is.numeric(key_work_wellness$Barrier_Mid_Managment_Support)

is.character(key_work_wellness$Barrier_Legal_Concerns)
suppressWarnings(key_work_wellness$Barrier_Legal_Concerns <- as.numeric(key_work_wellness$Barrier_Legal_Concerns))
is.numeric(key_work_wellness$Barrier_Legal_Concerns)

is.character(key_work_wellness$Barrier_Physical_Space)
suppressWarnings(key_work_wellness$Barrier_Physical_Space <- as.numeric(key_work_wellness$Barrier_Physical_Space))
is.numeric(key_work_wellness$Barrier_Physical_Space)

is.character(key_work_wellness$Incentive_Effectiveness)
suppressWarnings(key_work_wellness$Incentive_Effectiveness <- as.numeric(key_work_wellness$Incentive_Effectiveness))
is.numeric(key_work_wellness$Incentive_Effectiveness)

as_tibble(key_work_wellness)

summary(key_work_wellness)

```

## Assign Value Labels
Next, I assigned labels to the values of the character variables. I know there is different nomenclature in R, but my native statistical coding language is SPSS, so I'm defaulting to that. Again, if there is a more parsimonious way to code this, please let me know!

```{r}
key_work_wellness <-  key_work_wellness %>% 
    mutate(Your_Dept = recode(Your_Dept
                              ,"1" = "Human Resources and/or Benefits"
                              ,"2" = "Health Promotion/Fitness/Wellness"
                              ,"3" = "Safety"
                              ,"4" = "Medical"
                              ,"5" = "PR/Marketing"
                              ,"6" = "Finance"
                              ,"7" = "Other"))

key_work_wellness <-  key_work_wellness %>% 
    mutate(Region = recode(Region
                              ,"1" = "Northeast"
                              ,"2" = "Southeast"
                              ,"3" = "Midwest"
                              ,"4" = "Mountain"
                              ,"5" = "West"))

key_work_wellness <-  key_work_wellness %>% 
mutate(Industry = recode(Industry
                        ,"1" = "Agriculture, Forestry, Fishing, and Hunting; Mining; Utilities; Construction; Manufacturing"
                        ,"2" = "Wholesale Trade; Retail Trade; Transportation and Warehousing"
                        ,"3" = "Arts, Entertainment, and Recreation; Accommodation and Food Services; Other Services"
                        ,"4" = "Information; Finance' Insurance; Real Estate; Professional, Scientific, and Technical Services; Waste Management"
                        ,"5" = "Educational Services; Health Care and Social Assistance [excluding hospitals]"
                        ,"6" = "Public Administration"
                        ,"7" = "Hospital"))

key_work_wellness <-  key_work_wellness %>% 
    mutate(Num_Employees = recode(Num_Employees
                         ,"1" = "10 to 24"
                         ,"2" = "25 to 49"
                         ,"3" = "50 to 99"
                         ,"4" = "100 to 249"
                         ,"5" = "250 to 499"
                         ,"6" = "500 to 749"
                         ,"7" = "750 to 999"
                         ,"8" = "1000 or more")) 

key_work_wellness <-  key_work_wellness %>% 
    mutate(CDC_Region = recode(CDC_Region
                                  ,"1" = "CT, ME, MA, NH, RI, VT"
                                  ,"2" = "NJ, NY"
                                  ,"3" = "DE, DC, MD, PA, VA, WV"
                                  ,"4" = "AL, FL, GA, KY, MI, NC, SC, TN"
                                  ,"5" = "IL, IN, MI, MN, OH, WI"
                                  ,"6" = "AK, LA, NM, OK, TX"
                                  ,"7" = "IA, KS, MS, NB"
                                  ,"8" = "CO, MT, ND, SD, UT, WY"
                                  ,"9" = "AZ, CA, HI, NV"
                                  ,"10" = "AK, ID, OR, WA")) 

key_work_wellness <-  key_work_wellness %>% 
    mutate(Org_Type = recode(Org_Type
                               ,"1" = "For profit, public"
                               ,"2" = "For profit, private"
                               ,"3" = "Non-profit"
                               ,"4" = "State or local government"
                               ,"5" = "Federal government"
                               ,"6" = "Other")) 

key_work_wellness <-  key_work_wellness %>% 
    mutate(Stress_Manage = recode(Stress_Manage
                             ,"1" = "Yes"
                             ,"2" = "No")) 

key_work_wellness <-  key_work_wellness %>% 
    mutate(Stress_Manage_Type = recode(Stress_Manage_Type
                                  ,"1" = "Awareness or informational"
                                  ,"2" = "Skill-building"
                                  ,"3" = "Both informational and skill-building")) 

key_work_wellness <-  key_work_wellness %>% 
    mutate(Stress_Manage_Who_Runs = recode(Stress_Manage_Who_Runs
                                       ,"1" = "Offered mostly by employer"
                                       ,"2" = "Offered mostly by health plan"
                                       ,"3" = "Offered mostly by outsourced vendor"
                                       ,"4" = "Offered by employer, health plan and/or vendor")) 

key_work_wellness <-  key_work_wellness %>% 
    mutate(Stress_Manage_Particip = recode(Stress_Manage_Particip
                                           ,"1" = "1-25%"
                                           ,"2" = "26-50%"
                                           ,"3" = "51-75%"
                                           ,"4" = "More than 75%")) 

key_work_wellness <-  key_work_wellness %>% 
    mutate(Incentive_Effectiveness = recode(Incentive_Effectiveness
                                           ,"1" = "Not at all effective"
                                           ,"2" = "Somewhat effective"
                                           ,"3" = "Effective"
                                           ,"4" = "Extremely effective"))

key_work_wellness <-  key_work_wellness %>% 
    mutate(Health_Skill_Dev = recode(Health_Skill_Dev
                                     ,"1" = "Yes"
                                     ,"2" = "No")) 

key_work_wellness <-  key_work_wellness %>% 
    mutate(Supportive_Environment = recode(Supportive_Environment
                                     ,"1" = "Yes"
                                     ,"2" = "No")) 

key_work_wellness <-  key_work_wellness %>% 
    mutate(Integrate_Into_Org_Strucutre = recode(Integrate_Into_Org_Strucutre
                                     ,"1" = "Yes"
                                     ,"2" = "No")) 

key_work_wellness <-  key_work_wellness %>% 
    mutate(Bullying_Prevent = recode(Bullying_Prevent
                                  ,"1" = "Yes"
                                  ,"2" = "No")) 

key_work_wellness <-  key_work_wellness %>% 
    mutate(Flexible_Sched = recode(Flexible_Sched
                                     ,"1" = "Yes"
                                     ,"2" = "No")) 

key_work_wellness <-  key_work_wellness %>% 
    mutate(Telework = recode(Telework
                                   ,"1" = "Yes"
                                   ,"2" = "No")) 

key_work_wellness <-  key_work_wellness %>% 
    mutate(Family_Leave = recode(Family_Leave
                             ,"1" = "Yes"
                             ,"2" = "No")) 

key_work_wellness <-  key_work_wellness %>% 
    mutate(Parental_Leave_Paid = recode(Parental_Leave_Paid
                                 ,"1" = "Yes"
                                 ,"2" = "No")) 

key_work_wellness <-  key_work_wellness %>% 
    mutate(Parental_Leave_Unpaid = recode(Parental_Leave_Unpaid
                                        ,"1" = "Yes"
                                        ,"2" = "No")) 

key_work_wellness <-  key_work_wellness %>% 
    mutate(Childcare_Reimburse = recode(Childcare_Reimburse
                                          ,"1" = "Yes"
                                          ,"2" = "No")) 
key_work_wellness <-  key_work_wellness %>% 
    mutate(Childcare_OnOrOffsite = recode(Childcare_OnOrOffsite
                                        ,"1" = "Yes"
                                        ,"2" = "No")) 

key_work_wellness <-  key_work_wellness %>% 
    mutate(Edlercare_Reimburse = recode(Edlercare_Reimburse
                                          ,"1" = "Yes"
                                          ,"2" = "No")) 

key_work_wellness <-  key_work_wellness %>% 
    mutate(Eldercare_Onsite = recode(Eldercare_Onsite
                                        ,"1" = "Yes"
                                        ,"2" = "No")) 

key_work_wellness <-  key_work_wellness %>% 
    mutate(Incentive_Attendance = recode(Incentive_Attendance
                                     ,"1" = "Yes"
                                     ,"2" = "No")) 

key_work_wellness <-  key_work_wellness %>% 
    mutate(Incentive_Achieve_Health_Std = recode(Incentive_Achieve_Health_Std
                                         ,"1" = "Yes"
                                         ,"2" = "No")) 

key_work_wellness <-  key_work_wellness %>% 
    mutate(Incentive_Attendance_Health_Std = recode(Incentive_Attendance_Health_Std
                                                 ,"1" = "Yes"
                                                 ,"2" = "No")) 

key_work_wellness <-  key_work_wellness %>% 
    mutate(Incentive_Premium_Discounts = recode(Incentive_Premium_Discounts
                                                    ,"1" = "Yes"
                                                    ,"2" = "No")) 

key_work_wellness <-  key_work_wellness %>% 
    mutate(Incentive_Copay_Diff = recode(Incentive_Copay_Diff
                                                ,"1" = "Yes"
                                                ,"2" = "No")) 

key_work_wellness <-  key_work_wellness %>% 
    mutate(Incentive_Cash_Rewards = recode(Incentive_Cash_Rewards
                                         ,"1" = "Yes"
                                         ,"2" = "No")) 

key_work_wellness <-  key_work_wellness %>% 
    mutate(Incentive_Flex_Spending = recode(Incentive_Flex_Spending
                                           ,"1" = "Yes"
                                           ,"2" = "No")) 

key_work_wellness <-  key_work_wellness %>% 
    mutate(Incentive_Paid_Leave = recode(Incentive_Paid_Leave
                                            ,"1" = "Yes"
                                            ,"2" = "No")) 

key_work_wellness <-  key_work_wellness %>% 
    mutate(Incentive_Gifts_Prizes = recode(Incentive_Gifts_Prizes
                                         ,"1" = "Yes"
                                         ,"2" = "No")) 

key_work_wellness <-  key_work_wellness %>% 
    mutate(Yoga_Meditation_Room = recode(Yoga_Meditation_Room
                                           ,"1" = "Yes"
                                           ,"2" = "No")) 

head(key_work_wellness)
```

## Run tables for each column to make sure everything is looking good and lining up with the Codebook tabs
Again, please let me know if there is a more parsimonious way to do this!

```{r Run tables for each variable}
table(key_work_wellness$Industry)
table(key_work_wellness$Num_Employees)
table(key_work_wellness$CDC_Region)
table(key_work_wellness$Org_Type)
table(key_work_wellness$Yoga_Meditation_Room)
table(key_work_wellness$Stress_Manage)
table(key_work_wellness$Stress_Manage_Type)
table(key_work_wellness$Stress_Manage_Who_Runs)
table(key_work_wellness$Stress_Manage_Particip)
table(key_work_wellness$Percent_Under_30)
table(key_work_wellness$Percent_60_Or_Over)
table(key_work_wellness$Percent_Female)
table(key_work_wellness$Percent_Hourly)
table(key_work_wellness$Percent_Night_Shift)
table(key_work_wellness$Percent_Remote)
table(key_work_wellness$Percent_Turnover)
table(key_work_wellness$Bullying_Prevent)
table(key_work_wellness$Flexible_Sched)
table(key_work_wellness$Telework)
table(key_work_wellness$Family_Leave)
table(key_work_wellness$Parental_Leave_Paid)
table(key_work_wellness$Parental_Leave_Unpaid)
table(key_work_wellness$Childcare_Reimburse)
table(key_work_wellness$Childcare_OnOrOffsite)
table(key_work_wellness$Edlercare_Reimburse)
table(key_work_wellness$Eldercare_Onsite)
table(key_work_wellness$Barrier_Interest)
table(key_work_wellness$Barrier_Trust)
table(key_work_wellness$Barrier_Cost)
table(key_work_wellness$Barrier_Trained_Staff)
table(key_work_wellness$Barrier_Time)
table(key_work_wellness$Barrier_Privacy_Concerns)
table(key_work_wellness$Barrier_Outcomes)
table(key_work_wellness$Barrier_Qualified_Vendors)
table(key_work_wellness$Barrier_Sr_Managment_Support)
table(key_work_wellness$Barrier_Mid_Managment_Support)
table(key_work_wellness$Barrier_Legal_Concerns)
table(key_work_wellness$Barrier_Physical_Space)
table(key_work_wellness$Incentive_Attendance)
table(key_work_wellness$Incentive_Achieve_Health_Std)
table(key_work_wellness$Incentive_Attendance_Health_Std)
table(key_work_wellness$Incentive_Effectiveness)
table(key_work_wellness$Incentive_Premium_Discounts)
table(key_work_wellness$Incentive_Copay_Diff)
table(key_work_wellness$Incentive_Cash_Rewards)
table(key_work_wellness$Incentive_Flex_Spending)
table(key_work_wellness$Incentive_Paid_Leave)
table(key_work_wellness$Incentive_Gifts_Prizes)
```

## View of summary of the dataframe and check for missing values
```{r create a compact summary of the dataframe}
key_work_wellness %>%
  skim_without_charts()
```

## Save files as a CSV

```{r Save a copy of the data file to use in Google Sheets and Tableau}
write.csv(key_work_wellness, file = "key_work_wellness.csv")
```